---
title: "Cleaning_and_Wrangling_data"
author: "Zhe"
format: html
---
##Set up
```{r}
#| message: false
#| warning: false

library(dplyr)
library(tidyr)
library(readr)
```


## Read data
```{r}
#| warning: false
#| message: false
catch_orginal <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

## Explore Data
```{r}
colnames(catch_orginal)

head(catch_orginal)
summary(catch_orginal)
unique(catch_orginal$Region)

```

## Clean and Reshape


### Use 'select()' (by column)
```{r}
catch_data <- catch_orginal %>% 
  select(Region, Year, Sockeye, Chinook, Coho, Pink, Chum)
colnames(catch_data)
```

## Quality Check
```{r}
glimpse(catch_data)

```



### Replace values in column using 'mutate()'

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```

Finding rows with NA values
```{r}
i <- which(is.na(catch_clean$Chinook))
catch_data[i,]
```

Fixing I in Chinook column
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(condition = Chinook == "I",
                           true = "1",
                           false = Chinook),
         Chinook = as.numeric(Chinook))

head(catch_clean)
```

### Reshape data with 'pivot_longer()' and 'pivot_wider()'
```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year),
               names_to = "species",
               values_to = "catch")
head(catch_long)
```

Rerversion from long to wide
```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = species,
              values_from = catch)

head(catch_wide)
```

### Rename column using 'rename()'
```{r}
catch_long <- catch_long %>% 
  rename(catch_thousand = catch)
head(catch_long)
```

### Add a new column using 'mutate()'
```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch_thousand*1000) %>% 
  select(-catch_thousand)
head(catch_long)
```

### Demonstrate Split-Apply-Combine
```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))
head(mean_region)

n_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(n = n())

head(n_region)
``` 


### Use 'filter()' to subset by rows

```{r}
sse_catch <- catch_long %>% 
  filter(Region == "SSE")

unique(sse_catch$Region)
```

### Sort dataframe using 'arrange()'
```{r}



```


### Split and combine values in columnes using 'seperate ()' and 'unite()'
```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAU-039"))

sites_df %>%
    separate(site, c("island", "site_number"), "-")
```


```{r}
dates_df <- data.frame(
    year = c("1930",
             "1930",
             "1930"),
    month = c("12",
              "12",
              "12"),
    day = c("14",
            "15",
            "16")
)

dates_df %>%
    unite(date, year, month, day, sep = "-")
```
### All together using the pipe
```{r}
mean_region <- catch_orginal %>% 
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = if_else(Chinook == "I", 
                           "1",
                           Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year),
               names_to = "speicies",
               values_to = "catch") %>% 
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))

head(mean_region)

```




