---
title: "Quick analysis"
author: "Zhe"
format: html
editor: visual
---

# Introduction

## About the data

The data for this is donwloaded from the \[Arctic data center\] (https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X) on 10/07/2024.

## Setup

Attach important packges

```{r}
library(readr)
library(here)
```

## Read in the data

```{r}
#| message: false
#| warning: false
bg_chem <- read_csv(here('data', 'BGchem2008data.csv'))
```

# Analysis

## Calculate summary statistics

```{r}
#| eval: false
#| eco: false

### print the column names
colnames(bg_chem)
### get the general structure
str(bg_chem)
### First six line
head(bg_chem, 10)

### Get the summary of each column
summary(bg_chem)

### Unique values  for colume
unique(bg_chem$Date)

```

Cakcukaete the summary statistics (means) for nitrate ,nitrite, amonia, and ohisohate

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

## Calculate the mean Redfield ratio

Redfield ratio is the relative proportion of netrogen to phosphorus

```{r}
ratio <- (nitrate + nitrite + amm)/ phos
```

The redfield ratio for this dataset is approximately `r round(ratio)`

## Plot Redfield ratio

```{r}
#| echo: false


plot(bg_chem$P, bg_chem$NO3+ bg_chem$NO2+bg_chem$NH4)
```

# Conclusion
