---
title: "Tidy Data and Joins"
author: "Zhe"
format: html
---

### Read the libraray
```{r}
library(readr)
library(dplyr)
library(lubridate)
```
note: The following objects are masked from 'package:basse': data, inersect, setdiff,union.

### Read the data
```{r}
bird_observations <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=53edaa7a0e083013d9bf20322db1780e")
sites <- read.csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=81bf72420e69077097fb0790dcdc63a6")
surveys <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=b2466fa5cb5ed7ee1ea91398fc291c59")
taxalilst <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=58f863b7e3066e68536a9cacdc7bd58e")

```
### Get familiar with the data

```{r}
head(bird_observations)
dim(bird_observations)
glimpse(bird_observations)
summary(bird_observations)
summary(taxalilst)

```
### Create a subset of subset of bird_observations
```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>% 
  filter(site_id %in% c("LI-W", "NU-C"))
head(birds_subset)
```

### Use 'left_join()' to merge 'birds_subset' with the tabke 'sites'

```{r}
birds_left <- birds_subset %>% 
  left_join(y = sites, by = "site_id")
head(birds_left)
```

### Use 'full_join()' to merge 'birds_subset' and 'sites' tables 

```{r}
birds_full <- birds_subset %>% 
  full_join(y = sites, by = "site_id")
head(birds_full)
```

### Use 'inner_join()' to merge 'birds_subset' and 'taxalist' data
```{r}
birds_inner <- birds_subset %>% 
  inner_join(y = taxalilst, by = "species_id")

birds_left_taxa <- birds_subset %>% 
  left_join(y = taxalilst,by = "species_id" )
```

### Practice Wrangling & Joing data
```{r}
bird_observations <- bird_observations %>% 
  rename(bird_obs_notes = notes)
head(bird_observations)

birds_subset_v2 <- bird_observations %>% 
  full_join(y = taxalilst, by = "species_id") %>% 
  full_join(y = sites, by = "site_id") %>% 
  full_join(y = surveys, by = c("site_id", "survey_id")) %>% 
  select(survey_date, common_name, park_name, bird_count, observer)
```

### Explore 'oberver' data and fix the values within thic column so that all values are in the same format

```{r}
unique(birds_subset_v2$observer)
birds_subset_v2 <- birds_subset_v2 %>% 
  mutate(observer = if_else(condition = observer == "Josh Burns", 
                            true = "J. Burns", 
                            false = observer))

unique(birds_subset_v2$observer)
```

